### 2020-09-17 prf/thm-pcp-to-pgv-operational-correspondence-soundness.tex
Finished OC.

\begin{proof}
  \label{prf:thm-pcp-to-pgv-operational-correspondence-soundness}
  By induction on the derivation of $\pgv{\tm{\cpgvC{P}}\cred\tm{\conf{C}}}$.
  We omit the cases which cannot occur, \ie \LabTirName{E-New}, \LabTirName{E-Spawn}, and \LabTirName{LiftM}.

  \begin{case*}[\LabTirName{E-Link}]
    \[\pgv{%
        \tm{\res{x}{x'}{(\ppar{\plug{\conf{F}}{\link\;\pair{w}{x}}}{\conf{C}})}}
        \cred
        \tm{\ppar{\plug{\conf{F}}{\unit}}{\subst{\conf{C}}{w}{x'}}}
      }\]
    The source for $\pgv{\tm{\link\;\pair{w}{x}}}$ \emph{must} be $\pcp{\tm{\link{w}{x}}}$. None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be the empty context. Let $\pcp{\tm{P}}$ be the source term for $\pgv{\tm{\conf{C}}}$, \ie $\pgv{\tm{\cpgvC{P}}=\tm{\conf{C}}}$. Hence, we have:
    \begin{mathpar}
      \begin{tikzcd}[cramped, column sep=tiny]
        \pcp{\tm{\res{x}{x'}{(\ppar{\link{w}{x}}{P})}}}
        \arrow[r, "\pcp{\red}"]
        \arrow[d, "\cpgvC{\cdot}"]
        &
        \pcp{\tm{\subst{P}{w}{x'}}}
        \arrow[dd, "\cpgvC{\cdot}"]
        \\
        \pgv{\tm{\res{x}{x'}{(\ppar{\child\;\link\;\pair{w}{x}}{\cpgvC{P}})}}}
        \arrow[d, "\pgv{\cred^+}"]
        \\
        \pgv{\tm{\subst{\cpgvC{P}}{w}{x'}}}
        \arrow[r, "\pcp{=}"]
        &
        \pgv{\tm{\cpgvC{\subst{P}{w}{x'}}}}
      \end{tikzcd}
    \end{mathpar}
  \end{case*}
  \begin{case*}[\LabTirName{E-Send}]
    \[\pgv{%
        \tm{\res{x}{x'}{(\ppar{\plug{\conf{F}}{\send\;{\pair{V}{x}}}}{\plug{\conf{F'}}{\recv\;{x'}}})}}
        \cred
        \tm{\res{x}{x'}{(\ppar{\plug{\conf{F}}{x}}{\plug{\conf{F'}}{\pair{V}{x'}}})}}
      }\]
    There are three possible sources for $\pgv{\tm{\send}}$ and $\pgv{\tm{\recv}}$: $\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x'}{y'}{Q}}}$; $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\offer{x'}{Q}{R}}}$; or $\pcp{\tm{\inr{x}{P}}}$ and $\pcp{\tm{\offer{x'}{Q}{R}}}$.
    \begin{subcase*}[$\pcp{\tm{\send{x}{y}{P}}}$ and $\pcp{\tm{\recv{x'}{y'}{Q}}}$]
      None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be $\pgv{\tm{\child\;\letbind{x}{\hole}{\cpgvM{P}}}}$. Similarly, $\pgv{\tm{\conf{F'}}}$ must be $\pgv{\tm{\child\;\letpair{y'}{x'}{\hole}{\cpgvM{Q}}}}$. The value $\pgv{\tm{V}}$ must be an endpoint $\tm{y}$, bound by the name restriction $\pgv{\tm{\res{y}{y'}{}}}$ introduced by the translation. Hence, we have:
      \begin{mathpar}
          \begin{tikzcd}[cramped, column sep=tiny]
            \pcp{\tm{\res{x}{x'}{(\ppar{\send{x}{y}{P}}{\recv{x'}{y'}{Q}})}}}
            \arrow[d, "\cpgvC{\cdot}"]
            \arrow[r, "\pcp{\red}"]
            &
            \pcp{\tm{\res{x}{x'}{\res{y}{y'}{(\ppar{P}{Q})}}}}
            \arrow[dd, "\cpgvC{\cdot}"]
            \\
            \pgv{\tm{\res{x}{x'}{\res{y}{y'}{}}\left(
                  \begin{array}{l}
                    \child\;\letbind{x}{\send\;{\pair{y}{x}}}{\cpgvM{P}}
                    \parallel
                    \\
                    \child\;\letpair{y'}{x'}{\recv\;{x'}}{\cpgvM{Q}}
                  \end{array}
                \right)}}
            \arrow[d, "\pgv{\equiv\cred^+}"]
            \\
            \pgv{\tm{\res{x}{x'}{\res{y}{y'}{(\ppar{\child\;\cpgvM{P}}{\child\;\cpgvM{Q}})}}}}
            \arrow[r, "\pgv{\cred^\star}", "\text{(by \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC})}"']
            &
            \pgv{\tm{\res{x}{x'}{\res{y}{y'}{(\ppar{\cpgvC{P}}{\cpgvC{Q}})}}}}
          \end{tikzcd}
      \end{mathpar}
    \end{subcase*}
    \begin{subcase*}[$\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\offer{x'}{Q}{R}}}$]
      None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be $$\pgv{\tm{\child\;\letbind{x}{\andthen{\close\;\hole}{y}}{\cpgvM{P}}}}.$$ Similarly, $\pgv{\tm{\conf{F'}}}$ must be $$\pgv{\tm{\child\;\letpair{y'}{x'}{\hole}{\andthen{\wait\;x'}{\casesum{y'}{y'}{\cpgvM{Q}}{y'}{\cpgvM{R}}}}}}.$$ Hence, we have:
      \begin{mathpar}
        \begin{tikzcd}[cramped, column sep=tiny]
          \pcp{\tm{\res{x}{x'}{(\ppar{\inl{x}{P}}{\offer{x}{Q}{R}})}}}
          \arrow[d, "\cpgvM{\cdot}"]
          \arrow[r, "\pcp{\red}"]
          &
          \pcp{\tm{\res{x}{x'}{(\ppar{P}{Q})}}}
          \arrow[dd, "\cpgvC{\cdot}"]
          \\
          \pgv{\tm{\res{x}{x'}{}\left(
                \begin{array}{l}
                  \child\;\letbind{x}{\select{\labinl}\;{x}}{\cpgvM{P}}
                  \parallel
                  \\
                  \child\;\offer{x'}{x'}{\cpgvM{Q}}{x'}{\cpgvM{R}}
                \end{array}
              \right)}}
          \arrow[d, "\pgv{\cred^+}"]
          \\ 
          \pgv{\tm{\res{x}{x'}{(\ppar{\child\;\cpgvM{P}}{\child\;\cpgvM{Q}})}}}
          \arrow[r, "\pgv{\cred^\star}", "\text{(by \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC})}"']
          & 
          \pgv{\tm{\res{x}{x'}{(\ppar{\cpgvC{P}}{\cpgvC{Q}})}}}
        \end{tikzcd}
      \end{mathpar}
    \end{subcase*}
    \begin{subcase*}[$\pcp{\tm{\inr{x}{P}}}$ and $\pcp{\tm{\offer{x'}{Q}{R}}}$]
      None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be $$\pgv{\tm{\child\;\letbind{x}{\andthen{\close\;\hole}{y}}{\cpgvM{P}}}}.$$ Similarly, $\pgv{\tm{\conf{F'}}}$ must be $$\pgv{\tm{\child\;\letpair{y'}{x'}{\hole}{\andthen{\wait\;x'}{\casesum{y'}{y'}{\cpgvM{Q}}{y'}{\cpgvM{R}}}}}}.$$ Hence, we have:
      \begin{mathpar}
        \begin{tikzcd}[cramped, column sep=tiny]
          \pcp{\tm{\res{x}{x'}{(\ppar{\inr{x}{P}}{\offer{x}{Q}{R}})}}}
          \arrow[d, "\cpgvM{\cdot}"]
          \arrow[r, "\pcp{\red}"]
          &
          \pcp{\tm{\res{x}{x'}{(\ppar{P}{Q})}}}
          \arrow[dd, "\cpgvC{\cdot}"]
          \\
          \pgv{\tm{\res{x}{x'}{}\left(
                \begin{array}{l}
                  \child\;\letbind{x}{\select{\labinr}\;{x}}{\cpgvM{P}}
                  \parallel
                  \\
                  \child\;\offer{x'}{x'}{\cpgvM{Q}}{x'}{\cpgvM{R}}
                \end{array}
              \right)}}
          \arrow[d, "\pgv{\cred^+}"]
          \\ 
          \pgv{\tm{\res{x}{x'}{(\ppar{\child\;\cpgvM{P}}{\child\;\cpgvM{R}})}}}
          \arrow[r, "\pgv{\cred^\star}", "\text{(by \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC})}"']
          & 
          \pgv{\tm{\res{x}{x'}{(\ppar{\cpgvC{P}}{\cpgvC{R}})}}}
        \end{tikzcd}
      \end{mathpar}
    \end{subcase*}
  \end{case*}
  \begin{case*}[\LabTirName{E-Close}]
    \[\pgv{%
        \tm{\res{x}{x'}{(\ppar{\plug{\conf{F}}{\wait\;{x}}}{\plug{\conf{F'}}{\close\;{x'}}})}}
        \cred
        \tm{\ppar{\plug{\conf{F}}{\unit}}{\plug{\conf{F'}}{\unit}}}
      }\]
    The source for $\pgv{\tm{\wait}}$ and $\pgv{\tm{\close}}$ \emph{must} be $\pcp{\tm{\wait{x}{P}}}$ and $\pcp{\tm{\close{x'}{Q}}}$.

    (The translation for $\pcp{\tm{\offer{x}{P}{Q}}}$ also introduces a $\pgv{\tm{\wait}}$, but it is blocked on another communication, and hence cannot be the first communication on a translated term. The translations for $\pcp{\tm{\inl{x}{P}}}$ and $\pcp{\tm{\inr{x}{P}}}$ also introduce a $\pgv{\tm{\close}}$, but these are similarly blocked.)

    None of the translation rules introduce an evaluation context around the recursive call, hence $\pgv{\tm{\conf{F}}}$ must be $\pgv{\tm{\andthen{\hole}{\cpgvM{P}}}}$. Similarly, $\pgv{\tm{\conf{F'}}}$ must be $\pgv{\tm{\andthen{\hole}{\cpgvM{Q}}}}$. Hence, we have:
    \begin{mathpar}
      \begin{tikzcd}
        \pcp{\tm{\res{x}{x'}{(\ppar{\close{x}{P}}{\wait{x'}{Q}})}}} 
        \arrow[d, "\cpgvM{\cdot}"]
        \arrow[r, "\pcp{\red}"]
        &
        \pcp{\tm{\ppar{P}{Q}}}
        \arrow[dd, "\cpgvC{\cdot}"]
        \\
        \pgv{\tm{\res{x}{x'}{(\ppar
              {\child\;\andthen{\close\;{x}}{\cpgvM{P}}} 
              {\child\;\andthen{\wait\;{x'}}{\cpgvM{Q}}}
              )}}}
        \arrow[d, "\pgv{\cred^+}"]
        \\ 
        \pgv{\tm{\ppar{\child\;\cpgvM{P}}{\child\;\cpgvM{Q}}}}
        \arrow[r, "\pgv{\cred^\star}", "\text{(by \cref{lem:pcp-to-pgv-cpgvM-to-cpgvC})}"']
        & 
        \pgv{\tm{\ppar{\cpgvC{P}}{\cpgvC{Q}}}}
      \end{tikzcd}
    \end{mathpar}
  \end{case*}
  \begin{case*}[\LabTirName{E-LiftC}]
    By the induction hypothesis and \LabTirName{E-LiftC}.
  \end{case*}
  \begin{case*}[\LabTirName{E-LiftSC}]
    By the induction hypothesis, \LabTirName{E-LiftSC},
    and \cref{lem:pcp-to-pgv-confs-operational-correspondence-equiv}.
  \end{case*}
\end{proof}

%%% Local Variables:
%%% TeX-master: "../priorities"
%%% End: