### 2020-12-09 rebuttal.txt
Editted the rebuttal

We thank the reviewers for their thorough reviews and very useful feedback, which we'll address as follows.

First of all, we acknowledge the lack of examples and we commit to adding them to the potential final version.
We will illustrate PGV by adding examples for each of the following:

  - deadlock-free processes—including the streamlined version with two processes connected by two sessions, suggested by Reviewer 1—together with their typing derivations;
  - deadlocked processes that cannot be typed, thus illustrating how priorities work, suggested by Reviewers 1 and 2;
  - processes which use closures, thus illustrating the relevance of annotations on arrow types, suggested by Reviewer 1;
  - processes demonstrating the interaction between higher order and deadlock freedom, suggested by Reviewer 2.

Regarding the cyclic scheduler, we included it to deliberately connect our work to the previous FoSSaCS work on PCP.
We wanted to show how the well-known Milner's scheduler can be written in PGV (for the first time) and use it to illustrate our encoding of PCP to PGV, thus closing this circle on the cyclic scheduler.
We will make this clearer in the paper.

Regarding our contributions, we will highlight them better in the introduction. Summing them up here:
  - PGV is the first session-typed functional language with priorities *and* strong ties to linear logic via its correspondence with PCP and its relation to the CP and GV languages.
  - PGV addresses several problems in the original GV language, most notably:
    * our version does not have the pseudo-type S^#; and
    * our structural congruence is type preserving.

Lastly, we will implement all minor comments by all reviewers, thanks for pointing them out.

----------------------- REVIEW 1 ---------------------
Regarding the overlap with [37] we would like to mention that our work was developed completely independently of [37] as the authors came to know this work only during the writing up of the paper while working on related works.
We still want to emphasise the differences between our work and [37]:

  - PGV has a clear correspondence with a process calculus, PCP;
  - PGV has strong ties to linear logic, via its correspondence to PCP and its transitive relation to CP;

We will state the above clearly in the paper.

> As presented, Example 1 (also the PCP version) is ill-typed and, I think, not self contained.
> Note that d is used while d' is not and there is a restriction for d and d'.
> I think a `wait d'` is missing somewhere, possibly next to proc_1 if I interpret the diagram in the Appendix correctly.

Regarding M_i and N_i in the PCP cyclic scheduler, the reviewer understood this correctly and we will simply state that what matters is the communication structure being modelled. The missing `wait d'` should be in N_1, as the reviewer correctly inferred. We will mention this clearly in the paper.

Regarding the encoding of PCP to PGV and presenting only the encoding to the functional fragment, we have thought about this too. Such an encoding is complete but not sound, as the translation to configurations does some evaluation of top-level new/spawn constructs. For instance, the process ‘x[y].P’ has no reductions. Similarly, its translation to a configuration has no reductions. However, its translation to a term, ‘o let (y,z) = new in let x = send (z,x) in ((P))_M’ can reduce to ‘(vyz)(o let x = send (z,x) in ((P))_M)’.

Minor:

Regarding side conditions in SC-RESCOMM and E-NEW, the reviewer is right and we will add them.

Regarding T-VAR and T-CONST we will mention your suggestion.

----------------------- REVIEW 2 ---------------------

As mentioned at the top, we will add more examples to the paper.

Regarding the presentation of the main sections: we will explain more in detail how priorities work, especially with the aid of the aforementioned examples. We will systematise the the prevention of CP, PCP, GV and PGV, by dedicating subsections to each of them before diving into the details of our work.

Regarding our contribution, we will make them clearer for non-experts by first illustrating the use and benefits of GV and then the benefits of splitting new and spawn in PGV and its ties to logic (this will relate to the CP,PCP,GV,PGV explanations above).

----------------------- REVIEW 3 ---------------------

We will clarify the contributions better, this was also suggested by other reviewers.

It is true that priorities are not new and they were already used in PCP, which is a process calculus ties to logic. However, PGV is the first functional calculus with priorities that also ties to logic as PCP.

We are in fact making this connection to linear logic explicit by defining a sound encoding of PCP to PGV. Furthermore, as a case study for PGV and the encoding, we deliberately chose Milner's cyclic scheduler presented in PCP previously. The reason for this choice is to show how a process in the literature of PCP can be translated using our encoding into PGV.

However, in addition to the cyclic scheduler, we will add the examples mentioned at the top of this letter to better illustrate PGV and its type system.